\subject Glossary of \TeX\ PrimitivesThis chapter gives the list of all primitives of\TeX. After each control sequencethe grammatical category of the command or parameteris given, plus a short description. For some  commandsthe syntax of their use is given.For parameters the class to which they belong is given.Commands that have no grammatical categoryin \TeXbook\ are denoted either`\gr{expandable command}' or`\gr{primitive command}' in this list.Grammatical terms such as \gr{equals} and \gr{optional space}are explained in Chapter~\ref[gramm].\glossinvent\item -      \gr{horizontal command}       Discretionary hyphen; this is      equivalent to \cs{discretionary}""\ver|{-}{}{}|.      Can be used to indicate hyphenatable points in a word. \item \char32\      \gr{horizontal command}       Control space.      Insert the same amount of space as a space token would \alt      if \cs{spacefactor}${}=1000$.\item \char47\      \gr{primitive command}       Italic correction: insert a kern specified by the      preceding character.      Each character has an italic correction, possibly zero,      specified in the \n{tfm} file.      For slanted fonts this compensates for overhang.\item above\gr{dimen}      \gr{generalized fraction command}       Fraction with specified bar width. \item abovedisplayshortskip      \gr{glue parameter}      Glue above a display if the line preceding the display was short.\item abovedisplayskip      \gr{glue parameter}      Glue above a display.\item abovewithdelims\gr{delim$_1$}\gr{delim$_2$}\gr{dimen}      \gr{generalized fraction command}      Generalized fraction with delimiters.\item accent\gr{8-bit number}\gr{optional assignments}\gr{character}      \gr{horizontal command}      Command to place accents on characters.\alt\item adjdemerits      \gr{integer parameter}      Penalty for adjacent not visually compatible lines.       Default~\n{10$\,$000} in plain \TeX.\item advance\gr{numeric variable}\gr{optional \n{by}}\gr{number}      \gr{arithmetic assignment}      Arithmetic command to increase or decrease a       \gr{numeric variable}, that is, \alt      a \gr{count variable}, \gr{dimen variable}, \gr{glue variable},       or \gr{muglue variable}.\item afterassignment\gr{token}      \gr{primitive command}      Save the next token for execution after the next assignment.      Only one token can be saved this way.\item aftergroup\gr{token}      \gr{primitive command}      Save the next token for insertion after the current group.      Several tokens can be saved this way.\item atop\gr{dimen}      \gr{generalized fraction command}      Place objects over one another.\item atopwithdelims\gr{delim$_1$}\gr{delim$_2$}      \gr{generalized fraction command}      Place objects over one another with delimiters.\item badness      \gr{internal integer}       (\TeX3 only)       Badness of the most recently constructed box.\item baselineskip      \gr{glue parameter}       The `ideal' baseline distance between neighbouring       boxes on a vertical list;  \n{12pt} in plain \TeX.\item batchmode      \gr{interaction mode assignment}      \TeX\ patches errors itself  \alt      and performs an emergency stop on serious errors       such as missing input files,      but no terminal output is generated.\item begingroup      \gr{primitive command}      Open a group that must be closed with \ver-\endgroup-.\item belowdisplayshortskip      \gr{glue parameter}      Glue below a display if the line preceding the display was short.\item belowdisplayskip      \gr{glue parameter}      Glue below a display.\item binoppenalty      \gr{integer parameter}      Penalty for breaking after a binary operator not enclosed in      a subformula.      Plain \TeX\ default:~\n{700}.\item botmark      \gr{expandable command}      The last mark on the current page.\item box\gr{8-bit number}      \gr{box}      Use a box register, emptying it. \item boxmaxdepth      \gr{dimen parameter}      Maximum allowed depth of boxes.      Default~\cs{maxdimen} in plain \TeX.\item brokenpenalty      \gr{integer parameter}      Additional penalty for breaking a page after a hyphenated line.       Default~\n{100} in plain \TeX.\item catcode\gr{8-bit number}      \gr{internal integer}; the control sequence itself      is a~\gr{codename}.      Access category codes.\item char\gr{number}      \gr{character}      Explicit denotation of a character to be typeset. \item chardef\gr{control sequence}\gr{equals}\gr{number}      \gr{shorthand definition}      Define a control sequence to be a synonym for      a~character code.\item cleaders      \gr{leaders}      As \ver=\leaders=, but with box leaders       any excess space is split into equal glue items \alt      before and after the leaders.\item closein\gr{4-bit number}      \gr{primitive command}      Close an input stream.\item closeout\gr{4-bit number}      \gr{primitive command}      Close an output stream.\item clubpenalty      \gr{integer parameter}      Additional penalty for breaking a page after the first line of a paragraph.       Default~\n{150} in plain \TeX.\item copy\gr{8-bit number}      \gr{box}      Use a box register and retain the contents. \item count\gr{8-bit number}      \gr{internal integer}; the control sequence itself      is a~\gr{register prefix}.      Access count registers. \item countdef\gr{control sequence}\gr{equals}\gr{8-bit number}      \gr{shorthand definition}; the control sequence      itself is a~\gr{registerdef}.      Define a control sequence to be a synonym for      a~\cs{count} register.\item cr      \gr{primitive command}      Terminate an alignment line.\item crcr      \gr{primitive command}      Terminate an alignment line if it has       not already been terminated by~\cs{cr}.\item csname      \gr{expandable command}      Start forming the name of a control sequence.      Has to be balanced with \cs{endcsname}.\item day      \gr{integer parameter}      The day of the current job.\item deadcycles      \gr{special integer}      Counter that keeps track of how many times       the output routine has been called without a \cs{shipout}       taking place.       If this number reaches \cs{maxdeadcycles} \TeX\      gives an error message. Plain \TeX\ default:~\n{25}.\item def      \gr{def} Start a macro definition.\item defaulthyphenchar      \gr{integer parameter}      Value of \cs{hyphenchar} when a font is loaded.      Default value in plain \TeX\ is~\ver>`\->.\item defaultskewchar      \gr{integer parameter}      Value of \cs{skewchar} when a font is loaded.      Default value in plain \TeX\ is~\n{-1}.\item delcode\gr{8-bit number}      \gr{internal integer}; the control sequence itself      is a~\gr{codename}.      Access the      code specifying how a character should be used as delimiter       after \cs{left} or \cs{right}.\item delimiter\gr{27-bit number}      \gr{math character}      Explicit denotation of a delimiter.\item delimiterfactor      \gr{integer parameter}      1000 times the part of a delimited formula that should be      covered by a delimiter.      Plain \TeX\ default:~\n{901}.\item delimitershortfall      \gr{integer parameter}      Size of the part of a delimited formula that is allowed       to go uncovered by a delimiter.      Plain \TeX\ default:~\n{5pt}.\item dimen\gr{8-bit number}      \gr{internal dimen}; the control sequence itself      is a~\gr{register prefix}.      Access dimen registers.\item dimendef\gr{control sequence}\gr{equals}\gr{8-bit number}      \gr{shorthand definition}; the control sequence      itself is a~\gr{registerdef}.      Define a control sequence to be a synonym for      a~\cs{dimen} register.\item discretionary\SerifFont\italic\lb pre-break\rb\lb post-break\rb\lb no-break\rb{}      \gr{horizontal command}      Specify the way a character sequence is split up at a line break.\item displayindent      \gr{dimen parameter}      Distance by which the box, in which the display       is centred, is indented owing to hanging indentation.      This value is set automatically for each display.\item displaylimits      \gr{primitive command}      Restore default placement for limits.\item displaystyle      \gr{primitive command}      Select the display style of math typesetting.\item displaywidowpenalty      \gr{integer parameter}      Additional penalty for breaking a page before the last line       above a display formula.       Default~\n{50} in plain \TeX.\item displaywidth      \gr{dimen parameter}      Width of the box in which the display is centred.      This value is set automatically for each display.\item divide\gr{numeric variable}\gr{optional \n{by}}\gr{number}      \gr{arithmetic assignment}      Arithmetic command to divide a \gr{numeric variable}      (see \cs{advance}).\item doublehyphendemerits      \gr{integer parameter}      Penalty for consecutive lines ending with a hyphen.       Default~\n{10$\,$000} in plain \TeX.\item dp\gr{8-bit number}      \gr{internal dimen}; the control sequence itself      is a~\gr{box dimension}.      Depth of the box in a box register. \item dump      \gr{vertical command}      Dump a format file; possible only in \IniTeX,       not allowed inside a group.\item edef      \gr{def}      Start a macro definition;       the replacement text is expanded at definition time.\item else      \gr{expandable command}      Select      \gr{false text} of a conditional       or default case of \cs{ifcase}.\item emergencystretch      \gr{dimen parameter}       (\TeX3 only)       Assumed extra stretchability in lines of a paragraph      in third pass of the line-breaking algorithm.\item end      \gr{vertical command}      End this run.\item endcsname      \gr{expandable command}      Delimit the name of a control sequence that was begun      with \cs{csname}.\item endgroup      \gr{primitive command}      End a group that was opened with \ver-\begingroup-.\item endinput      \gr{expandable command}      Terminate inputting the current file after the current line.\item endlinechar       \gr{integer parameter}      The character code of the end-of-line character       appended to input lines.      \IniTeX\ default:~\n{13}.\item eqno\gr{math mode material}\n{\char36\char36}      \gr{eqno}      Place a right equation number in a display formula.\item errhelp      \gr{token parameter}      Tokens that will be displayed if the user       asks for help after an \cs{err\-message}.\item errmessage\gr{general text}      \gr{primitive command}      Report an error and give the user opportunity to act.\item errorcontextlines      \gr{integer parameter}       (\TeX3 only)      Number of additional context lines shown in error messages.\item errorstopmode      \gr{interaction mode assignment}        Ask for user input on the occurrence of an error.\item escapechar       \gr{integer parameter}      Number of the character that is  used       when control sequences are being converted      into character tokens.      \IniTeX\ default:~\n{92}.\item everycr      \gr{token parameter}      Token list inserted after every \cs{cr} or non-redundant \cs{crcr}.\item everydisplay      \gr{token parameter}      Token list inserted at the start of a display.\item everyhbox      \gr{token parameter}      Token list inserted at the start of a horizontal box.\item everyjob      \gr{token parameter}      Token list inserted at the start of each job.\item everymath      \gr{token parameter}      Token list inserted at the start of non-display math.\item everypar      \gr{token parameter}      Token list inserted in front of paragraph text.\item everyvbox      \gr{token parameter}      Token list inserted at the start of a vertical box.\item exhyphenpenalty      \gr{integer parameter}      Penalty for breaking a horizontal line at a discretionary      in the special case where the prebreak text is empty.       Default~\n{50} in plain \TeX.\item expandafter       \gr{expandable command}      Take the next two tokens and place the       expansion  of the second after the first.\item fam      \gr{integer parameter}      The number of the current font family.\item fi      \gr{expandable command}      Closing delimiter for all conditionals.\item finalhyphendemerits      \gr{integer parameter}      Penalty added when the penultimate line of a       paragraph ends with a hyphen.       Plain \TeX\ default~\n{5000}.\item firstmark      \gr{expandable command}      The first mark on the current page.\item floatingpenalty      \gr{integer parameter}      Penalty amount added to \cs{insertpenalties} \alt      when an insertion is split.\item font\gr{control sequence}\gr{equals}\gr{file name}\gr{at clause}      \gr{simple assignment}      Associate a control sequence with a \n{tfm} file.      When used on its own, this control sequence is a \gr{font},      denoting the current font.\item fontdimen\gr{number}\gr{font}      \gr{internal dimen}      Access various parameters of fonts.\item fontname\gr{font}      \gr{primitive command}      The external name of a font.\item futurelet\gr{control sequence}\gr{token$_1$}\gr{token$_2$}      \gr{let assignment}      Assign the meaning of \gr{token$_2$} to the      \gr{control sequence}.\item gdef      \gr{def}      Synonym for \ver-\global\def-.\item global      \gr{prefix}      Make the next definition, arithmetic statement,      or assignment global.\item globaldefs      \gr{integer parameter}      Override \cs{global} specifications: a positive value of this      parameter makes all assignments global, a negative value      makes them local.\item halign\gr{box specification}\lb\gr{alignment material}\rb{}      \gr{vertical command}      Horizontal alignment.      Display alignment:      \disp\n{\$\$}\cs{halign}\gr{box specification}\lb\n{...}\rb      \gr{optional assignments}\n{\$\$}\dispstop\item hangafter      \gr{integer parameter}      If positive, this denotes the number of lines       before indenting starts;       if negative, its absolute value is the number       of indented lines starting with the first line of the paragraph.       The default value of~1 is restored after every paragraph.\item hangindent      \gr{dimen parameter}      If positive, this indicates indentation from the left margin;       if negative, this is the negative of the indentation       from the right margin.       The default value of~\n{0pt} is restored after every paragraph.\item hbadness      \gr{integer parameter}      Threshold below which \TeX\ does not report an underfull       or overfull  horizontal box.      Plain \TeX\ default:~\n{1000}.\item hbox\gr{box specification}\lb\gr{horizontal material}\rb      \gr{box}      Construct a horizontal box. \item hfil      \gr{horizontal command}      Horizontal skip equivalent to \ver-\hskip 0cm plus 1fil-.\item hfill      \gr{horizontal command}      Horizontal skip equivalent to \ver-\hskip 0cm plus 1fill-.\item hfilneg      \gr{horizontal command}      Horizontal skip equivalent to \ver-\hskip 0cm minus 1fil-.\item hfuzz      \gr{dimen parameter}      Excess size that \TeX\ tolerates before it considers        a horizontal box overfull.      Plain \TeX\ default:~\n{0.1pt}.\item hoffset      \gr{dimen parameter}      Distance by which the page is shifted to the right       of the reference point which is at one inch from      the left margin.\item holdinginserts      \gr{integer parameter}       (only \TeX3)       If this is positive, insertions are not placed in their boxes       when the \cs{output} tokens are inserted.\item hrule      \gr{vertical command}      Rule that spreads in horizontal direction.\item hsize      \gr{dimen parameter}      Line width used for text typesetting inside a vertical box.\item hskip\gr{glue}       \gr{horizontal command}      Insert in horizontal mode a glue item.\item hss      \gr{horizontal command}      Horizontal skip equivalent to \ver-\hskip 0cm plus 1fil minus 1fil-.\item ht\gr{8-bit number}      \gr{internal dimen}; the control sequence itself      is a~\gr{box dimension}.      Height of the box in a box register. \item hyphenation\gr{general text}      \gr{hyphenation assignment}      Define hyphenation exceptions for the current value of \cs{language}.\item hyphenchar\gr{font}      \gr{internal integer}      Number of the character  behind which a       \ver-\discretionary{}{}{}- is inserted.\item hyphenpenalty      \gr{integer parameter}      Penalty associated with break at a discretionary in the general case.       Default~\n{50} in plain \TeX.\item if\gr{token$_1$}\gr{token$_2$}      \gr{expandable command}      Test equality of character codes. \item ifcase\gr{number}\gr{case$_0$}\cs{or}\n{...}\cs{or}%           \gr{case$_n$}\cs{else}\gr{other cases}\cs{fi}      \gr{expandable command}      Enumerated case statement.\item ifcat\gr{token$_1$}\gr{token$_2$}      \gr{expandable command}      Test whether two characters have the same category code.\item ifdim\gr{dimen$_1$}\gr{relation}\gr{dimen$_2$}      \gr{expandable command}      Compare two dimensions. \item ifeof\gr{4-bit number}      \gr{expandable command}      Test whether a file has been fully read, or does not exist.\item iffalse      \gr{expandable command}      This test is always false.\item ifhbox\gr{8-bit number}      \gr{expandable command}      Test whether a box register contains a horizontal box.\item ifhmode      \gr{expandable command}      Test whether the current mode is (possibly restricted) horizontal mode.\item ifinner      \gr{expandable command}      Test whether the current mode is an internal mode.\item ifmmode      \gr{expandable command}      Test whether the current mode is (possibly display) math mode.\item ifnum\gr{number$_1$}\gr{relation}\gr{number$_2$}      \gr{expandable command}      Test relations between numbers.\item ifodd\gr{number}      \gr{expandable command}      Test whether a number is odd.\item iftrue      \gr{expandable command}      This test is always true.\item ifvbox\gr{8-bit number}      \gr{expandable command}      Test whether a box register contains a vertical box. \item ifvmode      \gr{expandable command}      Test whether the current mode is (possibly internal) vertical mode.\item ifvoid\gr{8-bit number}      \gr{expandable command}      Test whether a box register is empty.\item ifx\gr{token$_1$}\gr{token$_2$}      \gr{expandable command}      Test equality of macro expansion, or equality of character code and      category code.\item ignorespaces        \gr{primitive command}      Expands following tokens until something other      than a~\gr{space token} is found.\item immediate       \gr{primitive command}      Prefix to have output operations executed right away.\item indent      \gr{primitive command}      Switch to horizontal mode and insert box with width \cs{parindent}.      This command is automatically inserted before a      \gr{horizontal       command} in vertical mode.\item input\gr{file name}      \gr{expandable command}      Read a specified file as \TeX\ input. \item inputlineno      \gr{internal integer}      (\TeX3 only)       Number of the current input line.\item insert\gr{8-bit number}\lb\gr{vertical mode material}\rb      \gr{primitive command}      Start an insertion item.\item insertpenalties      \gr{special integer}      Total of penalties for split insertions.      Inside the output routine the number of held-over insertions.\item interlinepenalty      \gr{integer parameter}      Penalty for breaking a page between lines of a paragraph.       Default~\n{0} in plain \TeX.\item jobname      \gr{expandable command}      Name of the main \TeX\ file being processed.\item kern\gr{dimen}      \gr{kern}      Add a kern item of the specified      \gr{dimen} to the list;      this can be used both in horizontal and vertical mode.\item language      \gr{integer parameter}       (\TeX3 only)      Choose a set of hyphenation patterns and exceptions.\item lastbox      \gr{box}      Register containing the last element  added to the current list,       if this was a box.\item lastkern      \gr{internal dimen}      If the last item on the list was a kern, the size of this.\item lastpenalty      \gr{internal integer}      If the last item on the list was a penalty, the value of this.\item lastskip      \gr{internal glue} or \gr{internal muglue}.      If the last item on the list was a skip, the size of this.\item lccode\gr{8-bit number}      \gr{internal integer}; the control sequence itself      is a~\gr{codename}.      Access the      character code that is the lowercase variant of a given code.\item leaders\gr{box or rule}\gr{vertical/horizontal/mathematical skip}      \gr{leaders}      Fill a specified amount of space with a rule or copies of box.\item left      \gr{primitive command}      Use the following character as an open delimiter.\item lefthyphenmin      \gr{integer parameter}       (\TeX3 only)      Minimum number of characters before a hyphenation.\item leftskip      \gr{glue parameter}      Glue that is placed to the left of all lines.\item leqno\gr{math mode material}\n{\char36\char36}      \gr{eqno}      Place a left equation number in a display formula.\item let\gr{control sequence}\gr{equals}\gr{token}      \gr{let assignment}      Define a control sequence to a token, assign its meaning      if the token is a command or macro. \item limits      \gr{primitive command}      Place limits over and under a large operator.      This is the default position in display style.\item linepenalty      \gr{integer parameter}      Penalty value associated with each line break.       Default~\n{10} in plain \TeX.\item lineskip       \gr{glue parameter}      Glue added if distance between bottom and top of neighbouring boxes       is less than \cs{lineskiplimit}.      Default~\n{1pt} in plain \TeX.\item lineskiplimit      \gr{dimen parameter}      Distance to be maintained between the bottom and top of       neighbouring boxes on a vertical list.      Default~\n{0pt} in plain \TeX.\item long      \gr{prefix}      Indicate that the arguments of the macro to be defined        are allowed to contain \cs{par} tokens.\item looseness       \gr{integer parameter}      Number of lines by which this paragraph has to be made longer       (or, if negative, shorter) than it would be ideally.\item lower\gr{dimen}\gr{box}      \gr{primitive command}      Adjust vertical positioning of a box in horizontal mode. \item lowercase\gr{general text}      \gr{primitive command}      Convert the argument to its lowercase form.\item mag      \gr{integer parameter}      1000 times the magnification of the document.      Default \n{1000} in \IniTeX.\item mark\gr{general text}      \gr{primitive command}      Specify a mark text.\item mathaccent\gr{15-bit number}\gr{math field}      \gr{primitive command}      Place an accent in math mode.\item mathbin\gr{math field}      \gr{math atom}      Let the following \gr{math field} function       as a binary operation.\item mathchar\gr{15-bit number}      \gr{primitive command}      Explicit denotation of a mathematical character.\item mathchardef\gr{control sequence}\gr{equals}\gr{15-bit number}      \gr{shorthand definition}      Define a control sequence to be a synonym for      a~math character code.\item mathchoice\lb\SerifFont {\it D\/\rb\lb T\/\rb\lb S\/\rb\lb SS\/}\rb      \gr{primitive command}      Give four variants of a formula for the four styles      of math typesetting.      \item mathclose\gr{math field}      \gr{math atom}      Let the following \gr{math field} function      as a closing symbol.\item mathcode\gr{8-bit number}      \gr{internal integer}; the control sequence itself      is a~\gr{codename}.      Code of a character determining its treatment in math mode.\item mathinner\gr{math field}      \gr{math atom}      Let the following \gr{math field} function       as an inner formula.\item mathop\gr{math field}      \gr{math atom}      Let the following \gr{math field} function       as a large operator.\item mathopen\gr{math field}      \gr{math atom}      Let the following \gr{math field} function       as an opening symbol.\item mathord\gr{math field}      \gr{math atom}      Let the following \gr{math field} function       as an ordinary object.\item mathpunct\gr{math field}      \gr{math atom}      Let the following \gr{math field} function       as a punctuation symbol.\item mathrel\gr{math field}      \gr{math atom}      Let the following \gr{math field} function as a relation.\item mathsurround      \gr{dimen parameter}      Kern amount placed before and after in-line formulas.\item maxdeadcycles      \gr{integer parameter}      The maximum number of times that the output routine is allowed to      be called without a \cs{shipout} occurring.      \IniTeX\ default:~\n{25}.\item maxdepth      \gr{dimen parameter}      Maximum depth of the page box.      Default~\n{4pt} in plain \TeX.\item meaning      \gr{expandable command}      Give the meaning of a control sequence as a string of characters.\item medmuskip      \gr{muglue parameter}      Medium amount of mu glue.      Default value in plain \TeX: \n{4mu plus 2mu minus 4mu}\item message\gr{general text}      \gr{primitive command}      Write a message to the terminal.      \item mkern      \gr{primitive command}      Insert a kern measured in mu units.\item month      \gr{integer parameter}      The month of the current job.\item moveleft\gr{dimen}\gr{box}      \gr{primitive command}      Adjust horizontal positioning of a box in vertical mode. \item moveright\gr{dimen}\gr{box}      \gr{primitive command}      Adjust horizontal positioning of a box in vertical mode. \item mskip      \gr{mathematical skip}      Insert glue measured in mu units.\item multiply\gr{numeric variable}\gr{optional \n{by}}\gr{number}      \gr{arithmetic assignment}      Arithmetic command to multiply a      \gr{numeric variable} (see \cs{advance}).\item muskip\gr{8-bit number}      \gr{internal muglue}; the control sequence itself      is a~\gr{register prefix}.      Access skips measured in mu units. \item muskipdef\gr{control sequence}\gr{equals}\gr{8-bit number}      \gr{shorthand definition}; the control sequence      itself is a~\gr{registerdef}.      Define a control sequence to be a synonym for      a~\cs{muskip} register.\item newlinechar      \gr{integer parameter}      Number of the character that triggers a new line in      \cs{write} statements.      Plain \TeX\ default~$-1$.      \item noalign\gr{filler}\lb\gr{vertical (horizontal) %           mode material}\rb      \gr{primitive command}      Specify  vertical (horizontal)      material   to be placed in between rows (columns) of      an \cs{halign} (\cs{valign}).\item noboundary      \gr{horizontal command}      (\TeX3 only)      Omit implicit boundary character.\item noexpand\gr{token}      \gr{expandable command}      Do not expand the next token.\item noindent       \gr{primitive command}      Switch to horizontal mode with an empty horizontal list.\item nolimits      \gr{primitive command}      Place limits of a large operator as subscript and       superscript expressions.      This is the default position in text style.\item nonscript      \gr{primitive command}      Cancel the next glue item if it occurs in       scriptstyle or scriptscriptstyle.\item nonstopmode      \gr{interaction mode assignment}      \TeX\ fixes errors as best it can,      and performs an emergency stop      when user interaction is needed.\item nulldelimiterspace      \gr{dimen parameter}      Width taken for empty delimiters.       Default~\n{1.2pt} in plain \TeX.\item nullfont      \gr{fontdef token}      Name of an empty font that \TeX\ uses in emergencies.\item number\gr{number}      \gr{expandable command}      Convert a      \gr{number} to decimal representation. \item omit      \gr{primitive command}      Omit the template for one alignment entry.\item openin\gr{4-bit number}\gr{equals}\gr{filename}      \gr{primitive command}      Open a stream for input.\item openout\gr{4-bit number}\gr{equals}\gr{filename}      \gr{primitive command}      Open a stream for output.\item or      \gr{primitive command}      Separator for entries of an \cs{ifcase}.\item outer      \gr{prefix}      Indicate that the macro being defined       should occur on the outer level only.\item output      \gr{token parameter}      Token list with instructions for shipping out pages.\item outputpenalty        \gr{integer parameter}      Value of the penalty at the current page break,      or $10\,000$ if the break was not at a penalty.\item over      \gr{generalized fraction command}      Fraction.\item overfullrule      \gr{dimen parameter}      Width of the rule that is printed to indicate       overfull horizontal boxes.      Plain \TeX\ default:~\n{5pt}.\item overline\gr{math field}      \gr{math atom}      Overline the following \gr{math field}.\item overwithdelims\gr{delim$_1$}\gr{delim$_2$}      \gr{generalized fraction command}      Fraction with delimiters.\item pagedepth      \gr{special dimen}      Depth of the current page.\item pagefilllstretch      \gr{special dimen}      Accumulated third-order stretch of the current page.\item pagefillstretch      \gr{special dimen}      Accumulated second-order stretch of the current page.\item pagefilstretch      \gr{special dimen}      Accumulated first-order stretch of the current page.\item pagegoal      \gr{special dimen}      Goal height of the page box. This starts at \cs{vsize},      and is diminished by heights of insertion items.\item pageshrink      \gr{special dimen}      Accumulated shrink of the current page.\item pagestretch      \gr{special dimen}      Accumulated zeroth-order stretch of the current page.\item pagetotal      \gr{special dimen}      Accumulated natural height of the current page.\item par      \gr{primitive command}      Close off a paragraph and go into vertical mode.\item parfillskip      \gr{glue parameter}      Glue that is placed between the last                element of the paragraph and the line end.      Plain \TeX\ default:~\n{0pt plus 1fil}.\item parindent      \gr{dimen parameter}      Size of the indentation box added in front of a paragraph.\item parshape      \gr{internal integer}      Command for general paragraph shapes:       \disp\cs{parshape}\gr{equals}$n$ $i_1$ $\ell_1$ $\ldots$                $i_n$ $\ell_n$\>      specifies a~number      of lines~$n$, and $n$~pairs of an indentation and      line length.\item parskip      \gr{glue parameter}      Amount of glue added to vertical list when a paragraph starts;       default value \ver.0pt plus 1pt. in plain \TeX.\item patterns\gr{general text}      \gr{hyphenation assignment}      Define a list of hyphenation patterns for the current      value of \cs{language};  allowed only in \IniTeX.\item pausing      \gr{integer parameter}      Specify that \TeX\ should pause after each line that is       read from a file.\item penalty      \gr{primitive command}      Specify desirability of not breaking at this point.\item postdisplaypenalty      \gr{integer parameter}      Penalty placed in the vertical list below a display.\item predisplaypenalty      \gr{integer parameter}      Penalty placed in the vertical list above a display.      Plain \TeX\ default:~\n{10$\,$000}.\item predisplaysize      \gr{dimen parameter}      Effective width of the line preceding the display.\item pretolerance      \gr{integer parameter}      Tolerance value for a paragraph that uses no hyphenation.       Default~\n{100} in plain \TeX.\item prevdepth       \gr{special dimen}      Depth of the last box added to a vertical list as it is       perceived by \TeX.\item prevgraf       \gr{special integer}      The number of lines in the paragraph last      added to the vertical list.\item radical\gr{24-bit number}      \gr{primitive command}      Command for setting things such as root signs.\item raise\gr{dimen}\gr{box}      \gr{primitive command}      Adjust vertical positioning of a box in horizontal mode. \item read\gr{number}\n{to}\gr{control sequence}      \gr{simple assignment}      Read a line from a stream into a control sequence.\item relax      \gr{primitive command}      Do nothing.\item relpenalty      \gr{integer parameter}      Penalty for breaking after a binary relation, not enclosed      in a subformula.      Plain \TeX\ default:~\n{500}.\item right      \gr{primitive command}      Use the following character as a closing delimiter.\item righthyphenmin      \gr{integer parameter}       (\TeX3 only)       Minimum number of characters after a hyphenation.\item rightskip      \gr{glue parameter}      Glue that is placed to the right of all lines.\item romannumeral\gr{number}      \gr{expandable command}      Convert a positive      \gr{number} to lowercase roman representation.\item scriptfont\gr{4-bit number}      \gr{family member}; the control sequence itself      is a~\gr{font range}.      Access the scriptfont of a family.\item scriptscriptfont\gr{4-bit number}      \gr{family member}; the control sequence itself      is a~\gr{font range}.      Access the scriptscriptfont of a family.\item scriptscriptstyle      \gr{primitive command}      Select the scriptscript style of math typesetting.\item scriptspace      \gr{dimen parameter}      Extra space after subscripts and superscripts.      Default~\n{.5pt}      in plain \TeX.\item scriptstyle      \gr{primitive command}      Select the script style of math typesetting.\item scrollmode      \gr{interaction mode assignment}        \TeX\ patches errors itself, but will ask the user for missing files.\item setbox\gr{8-bit number}\gr{equals}\gr{box}      \gr{simple assignment}      Assign a box to a box register.\item setlanguage\gr{number}      \gr{primitive command}       (\TeX3 only)       Insert a whatsit resetting the current language      to the \gr{number} specified.\item sfcode\gr{8-bit number}      \gr{internal integer}; the control sequence itself      is a~\gr{codename}.      Access the value of the \cs{spacefactor}      associated with a character.\item shipout\gr{box}      \gr{primitive command}      Ship a box to the \n{dvi} file.\item show\gr{token}      \gr{primitive command}      Display the meaning of a token on the screen.\item showbox\gr{8-bit number}      \gr{primitive command}      Write the contents of a box to the log file.\item showboxbreadth      \gr{integer parameter}      Number of successive elements that are shown when       \ver-\tracingoutput- is positive, each time a level is visited.      Plain \TeX\ default:~\n{5}.\item showboxdepth       \gr{integer parameter}      The number of levels that are shown when       \ver-\tracingoutput- is positive.      Plain \TeX\ default:~\n{3}.\item showlists      \gr{primitive command}      Write to the log file the contents of the partial lists       currently built in all modes.\item showthe\gr{internal quantity}      \gr{primitive command}      Display on the terminal the result       of prefixing a token with \cs{the}.\item skewchar\gr{font}      \gr{internal integer}      Font position of an after-placed accent.      \item skip\gr{8-bit number}      \gr{internal glue}; the control sequence itself      is a~\gr{register prefix}.      Access skip registers\item skipdef\gr{control sequence}\gr{equals}\gr{8-bit number}      \gr{shorthand definition}; the control sequence      itself is a~\gr{registerdef}.      Define a control sequence to be a synonym for      a~\cs{skip} register.\item spacefactor      \gr{special integer}      1000 times the ratio by which the stretch component of the      interword glue should be multiplied.\item spaceskip      \gr{glue parameter}      Interword glue if non-zero.\item span      \gr{primitive command}      Join two adjacent alignment entries, or (in preamble)      expand the next token.\item special\gr{general text}      \gr{primitive command}      Write a token list  to the \n{dvi} file.\item splitbotmark      \gr{expandable command}      The last mark on a split-off page.\item splitfirstmark      \gr{expandable command}      The first mark on a split-off page.\item splitmaxdepth      \gr{dimen parameter}      Maximum depth of a box split off by a \cs{vsplit} operation.       Default~\n{4pt} in plain \TeX.\item splittopskip      \gr{glue parameter}      Minimum distance between the top of what remains after a      \cs{vsplit} operation, and the first item in that box.      Default~\n{10pt} in plain \TeX.\item string\gr{token}      \gr{expandable command}      Convert a token to a string of one or more characters. \item tabskip      \gr{glue parameter}      Amount of glue in between columns (rows) of an \cs{halign} \alt      (\cs{valign}).\item textfont\gr{4-bit number}      \gr{family member}; the control sequence itself      is a~\gr{font range}.      Access the textfont of a family.\item textstyle      \gr{primitive command}      Select the text style of math typesetting.\item the\gr{internal quantity}      \gr{primitive command}      Expand the value of various quantities in \TeX\ into a string      of (character) tokens.\item thickmuskip      \gr{muglue parameter}      Large amount of mu glue.       Default value in plain \TeX: \n{5mu plus 5mu}.\item thinmuskip      \gr{muglue parameter}      Small amount of mu glue.      Default value in plain \TeX: \n{3mu}.\item time      \gr{integer parameter}      Number of minutes after midnight that the current job started.\item toks\gr{8-bit number}      \gr{register prefix}      Access a token list register.\item toksdef\gr{control sequence}\gr{equals}\gr{8-bit number}      \gr{shorthand definition}; the control sequence      itself is a~\gr{registerdef}.      Assign a control sequence to      a~\cs{toks} register.\item tolerance      \gr{integer parameter}        Tolerance value for lines in a paragraph that does use hyphenation.       Default~\n{200} in plain \TeX, \n{10$\,$000} in \IniTeX.\item topmark      \gr{expandable command}      The last mark of the previous page.\item topskip      \gr{glue parameter}      Minimum distance between the top of the page box      and the baseline of the first box on the page.       Default~\n{10pt} in plain \TeX.\item tracingcommands      \gr{integer parameter}      When this is~1, \TeX\ displays primitive commands executed;       when this is 2~or more the outcome of conditionals is also recorded.\item tracinglostchars      \gr{integer parameter}            If this parameter is positive, \TeX\ gives            diagnostic messages whenever a character is accessed that            is not present in a font. Plain \TeX\ default:~1.\item tracingmacros      \gr{integer parameter}      If this is~1, the log file shows expansion of macros       that are performed and the actual values of the arguments;       if this is 2~or more      \gr{token parameter}s such as       \cs{output} and \cs{everypar} are also traced.\item tracingonline      \gr{integer parameter}            If this parameter is positive, \TeX\ will write trace            information also to the terminal.\item tracingoutput      \gr{integer parameter}      If this parameter is positive, the log file shows a dump of boxes       that are shipped to the \n{dvi} file.\item tracingpages      \gr{integer parameter}            If this parameter is positive, \TeX\ generates            a trace of the page-breaking algorithm.\item tracingparagraphs      \gr{integer parameter}            If this parameter is positive, \TeX\ generates            a trace of the line-breaking algorithm.\item tracingrestores      \gr{integer parameter}            If this parameter is positive, \TeX\ will report            all values that are restored when a group level ends.\item tracingstats      \gr{integer parameter}            If this parameter is positive, \TeX\ reports at the            end of the job the usage of various internal arrays.\item uccode\gr{8-bit number}      \gr{internal integer}; the control sequence itself      is a~\gr{codename}.      Access      the character code that is the uppercase variant of a given code.\item uchyph      \gr{integer parameter}      Positive if hyphenating words starting with a capital       letter is allowed.       Plain \TeX\ default~1.\item underline\gr{math field}      \gr{math atom}      Underline the following \gr{math field}.\item unhbox\gr{8-bit number}      \gr{horizontal command} \alt      Unpack a box register containing a horizontal box,        appending the contents to the list, and emptying the register. \item unhcopy\gr{8-bit number}      \gr{horizontal command} \alt      The same as \cs{unhbox}, but do not empty the register. \item unkern      \gr{primitive command}      Remove the last item of the list if this was a kern.\item unpenalty      \gr{primitive command}      Remove the last item of the list if this was a penalty.\item unskip      \gr{primitive command}      Remove the last item of the list if this was a skip.\item unvbox\gr{8-bit number}      \gr{vertical command} \alt      Unpack a box register containing a vertical box,       appending the contents to the list, and emptying the register. \item unvcopy\gr{8-bit number}      \gr{vertical command} \alt      The same as \cs{unvbox}, but do not empty the register. \item uppercase\gr{general text}      \gr{primitive command}      Convert the argument to its uppercase form.\item vadjust\gr{filler}\lb\gr{vertical mode material}\rb      \gr{primitive command}      Specify in horizontal mode material for the enclosing vertical list.\item valign\gr{box specification}\lb\gr{alignment material}\rb      \gr{horizontal command}         Vertical alignment.   \item vbadness      \gr{integer parameter}      Threshold below which overfull and underfull vertical boxes       are not shown.      Plain \TeX\ default:~\n{1000}.\item vbox\gr{box specification}\lb\gr{vertical material}\rb      \gr{primitive command}      Construct a vertical box with reference point on the last item. \item vcenter\gr{box specification}\lb\gr{vertical material}\rb      \gr{primitive command}      Construct a  vertical box vertically centred on the math axis.\item vfil      \gr{vertical command}      Vertical skip equivalent to \ver-\vskip 0cm plus 1fil-.\item vfill      \gr{vertical command}      Vertical skip equivalent to \ver-\vskip 0cm plus 1fill-.\item vfilneg      \gr{vertical command}      Vertical skip equivalent to \ver-\vskip 0cm minus 1fil-.\item vfuzz      \gr{dimen parameter}       Excess size that \TeX\ tolerates before it considers        a vertical box overfull.      Plain \TeX\ default:~\n{0.1pt}.\item voffset      \gr{dimen parameter}      Distance by which the page is shifted down from the reference point,      which is one inch from the top of the page.\item vrule      \gr{horizontal command}      Rule that spreads in vertical direction.\item vsize      \gr{dimen parameter}      Height of the page box.\item vskip\gr{glue}      \gr{vertical command}      Insert in vertical mode a glue item.\item vsplit\gr{8-bit number}\n{to}\gr{dimen}      \gr{primitive command}      Split off the top part of a vertical box. \item vss      \gr{vertical command}      Vertical skip equivalent to \ver-\vskip 0cm plus 1fil minus 1fil-.\item vtop\gr{box specification}\lb\gr{vertical material}\rb      \gr{primitive command}      Construct a vertical box with reference point on the first item. \item wd\gr{8-bit number}      \gr{internal dimen}; the control sequence itself      is a~\gr{box dimension}.      Width of the box in a box register.\item widowpenalty      \gr{integer parameter}      Additional penalty for breaking a page before       the last line of a paragraph.       Default~\n{150} in plain \TeX.\item write\gr{number}\gr{general text}      \gr{primitive command}      Generate a whatsit item containing      a~token list to be written to the terminal or to a file. \item xdef      \gr{def}      Synonym for \ver-\global\edef-.\item xleaders      \gr{leaders}      As \cs{leaders}, but with box leaders any excess space is       spread equally between the boxes.\item xspaceskip      \gr{glue parameter}      Interword glue if non-zero and \cs{spacefactor}${}\geq2000$.\item year      \gr{integer parameter}      The year of the current job.\>\endinput