\subject[token] Token Lists\TeX\ has only one type of data structure: the token list.\term token lists\parThere are token list registers that are available to the user,and \TeX\ has some special token lists: the \cs{every...}variables, \cs{errhelp}, and \cs{output}.\invent\item toks       Prefix for a token list register.\item toksdef       Define a control sequence to be a synonym for      a~\cs{toks} register.\item newtoks       Macro that allocates a token list register.\inventstop\point Token listsToken lists are the only type of data structure that \TeX\ knows.They can contain character tokens and control sequence tokens.Spaces in a token list are significant.The only operations on token lists are assignment andunpacking.\TeX\ has 256 token list registers \ver|\toks|$nnn$ that can beallocated using the macro \ver|\newtoks|, or explicitlyassigned by \cs{toksdef}; see below.\point Use of token listsToken lists are  assigned by a \gr{variable assignment},which is in this case takes one of the forms\disp\gr{token variable}\gr{equals}\gr{general text}\nl     \gr{token variable}\gr{equals}\gr{filler}\gr{token variable}\dispstopHere a \gr{token variable} is an explicit \cs{toks}$nnn$register, something that has been defined to such a registerby \cs{toksdef} (probably hidden in \cs{newtoks}),or one of the special \gr{token parameter}lists below.A~\gr{general text} has an explicit closing brace, but theopen brace can be implicit.Examples of token lists are (the first two lines are equivalent):\Ver>\toks0=\bgroup \a \b cd}\toks0={\a \b cd}\toks1=\toks2<RevUnpacking a token list is done by the command \cs{the}:the expansion of \cs{the}\gr{token variable} is the sequence of tokens that was in the token list.Token lists have a special behaviour in \cs{edef}:when prefixed by \ver|\the| they are unpacked, but the resulting tokensare not evaluated further. Thus\Ver>\toks0={\a \b} \edef\SomeCs{\the\toks0}<Revgives\Ver>\SomeCs: macro:-> \a \b<RevThis is in contrast to what happens ordinarily in an~\cs{edef};see page~\pgref[expand:edef].\point \gr{token parameter}There are in \TeX\ a number of token lists that are automaticallyinserted at certain points. These \gr{token parameter}s arethe following:\description \item \cs{output}   this token list is inserted   whenever \TeX\ decides it has sufficient material for a page,   or when the user forces activation by a penalty~$\leq-10\,000$   in vertical mode   (see Chapter~\ref[output]);\item \cs{everypar}   is inserted when \TeX\ switches from external or internal   vertical mode to unrestricted horizontal mode    (see Chapter~\ref[par:start]);\item \cs{everymath}   is inserted after a single math-shift character that starts   a formula;\item \cs{everydisplay}   is inserted after a double math-shift character that starts   a display formula;\item \cs{everyhbox}   is inserted when an \cs{hbox} begins (see Chapter~\ref[boxes]);\item \cs{everyvbox}   is inserted when a vertical box begins (see Chapter~\ref[boxes]);\item \cs{everyjob}   is inserted when a job begins (see Chapter~\ref[run]);\item \cs{everycr}   is inserted in alignments after \cs{cr} or a non-redundant   \cs{crcr} (see Chapter~\ref[align]);\item \cs{errhelp}   contains tokens to supplement an \cs{errmessage}    (see Chapter~\ref[error]).\descriptionstopA \gr{token parameter} behaves the same as an explicit \cs{toks}$nnn$list, or a quantity defined by \cs{toksdef}.\point Token list registersToken lists can be stored in \cs{toks} registers:   \csterm toks\par\Disp\cs{toks}\gr{8-bit number}\Dispstopwhich is a \gr{token variable}.Synonyms for token list registers can be made by the \gr{registerdef}command \cs{toksdef} in a \gr{shorthand definition}:   \csterm toksdef\par\Disp\cs{toksdef}\gr{control sequence}\gr{equals}\gr{8-bit number}\Dispstop A control sequence defined this way is calleda \gr{toksdef token}, and this is also a token variable(the remaining third kind of token variable isthe \gr{token parameter}).The plain \TeX\ macro \cs{newtoks} uses \cs{toksdef} to    \csterm newtoks\parallocate unused token list registers. This macro is \cs{outer}.\point ExamplesToken lists are probably among the least obvious componentsof \TeX: most \TeX\ users will never find occasion  for their use,but format designers and other macro writerscan find interesting applications.Following are some examples of the sorts of things that can bedone with token lists.\spoint Operations on token lists: stack macrosThe number of primitive operations available for token lists is\howto Stack macros\parrather limited: assignment and unpacking. However, these aresufficient to implement other operations such as appending.Let us say we have allocated a token register\Ver>\newtoks\list \list={\c} <Revand we want to add tokens to it,\altusing the syntax\Ver>\Prepend \a \b (to:)\list<Revsuch that \Ver>\showthe\list<Rev gives \Ver>> \a \b \c .<RevFor this the original list has to be unpacked, and \altthe new tokens followed by the old contents have to assignedagain to the register. Unpacking can be done with \cs{the}inside an \cs{edef}, so we arrive at the following macro:\Ver>\def\Prepend#1(to:)#2{\toks0={#1}%    \edef\act{\noexpand#2={\the\toks0 \the#2}}%    \act}<RevNote that the tokens that are to be added are first packed\altinto a temporary token list, which is then again unpackedinside the \cs{edef}. Including them directly would haveled to their expansion.Next we want to use token lists as a sort of stack:we want a `pop' operation that removes the first elementfrom the list. Specifically,\Ver>\Pop\list(into:)\first\show\first \showthe\list<Revshould give\Ver>> \first=macro:->\a .<Revand for the remaining list\Ver> > \b \c .<RevHere we make creative use of delimited and undelimitedparameters. With an \cs{edef} we unpack the list,and the auxiliary macro \cs{SplitOff} scoops up the elementsas one undelimited argument, the first element, and onedelimited argument, the rest of the elements.\Ver>\def\Pop#1(into:)#2{%    \edef\act{\noexpand\SplitOff\the#1%              (head:)\noexpand#2(tail:)\noexpand#1}%    \act}\def\SplitOff#1#2(head:)#3(tail:)#4{\def#3{#1}#4={#2}}<Rev\spoint Executing token listsThe \cs{the} operation for unpacking token lists was used aboveonly inside an \cs{edef}. Used on its own it has the effectof feeding the tokens of the list to \TeX's expansion mechanism.If the tokens have been added to the list in a uniform syntax,this gives rise to some interesting possibilities.Imagine that we are implementing the bookkeeping of externalfiles for a format. Such external files can be used fortable of contents, list of figures, et cetera.If the presenceof such objects is under the control of the user, we need somegeneral routines for opening and closing files, and keepingtrack of what files we have opened at the user's request.Here only  some routines for bookkeeping will be described.Let us say there is a list of auxiliary files, and an auxiliarycounter: \Ver>\newtoks\auxlist \newcount\auxcount<RevFirst of all there must be an operation to add auxiliary files:\Ver>\def\NewAuxFile#1{\AddToAuxList{#1}%    % plus other actions    }\def\AddToAuxList#1{\let\\=\relax    \edef\act{\noexpand\auxlist={\the\auxlist \\{#1}}}%    \act}<RevThis adds the name to the list in a uniform format:\Ver>\NewAuxFile{toc} \NewAuxFile{lof}\showthe\auxlist> \\{toc}\\{lof}.<Revusing the control sequence \ver>\\> which is left undefined.Now this control sequence can be used for instance tocount the number of elements in the list:\Ver>\def\ComputeLengthOfAuxList{\auxcount=0     \def\\##1{\advance\auxcount1\relax}%    \the\auxlist}\ComputeLengthOfAuxList \showthe\auxcount> 2.<RevAnother use of this structure is the following:at the end of the job we can now close all auxiliaryfiles at once, by\Ver>\def\CloseAuxFiles{\def\\##1{\CloseAuxFile{##1}}%    \the\auxlist}\def\CloseAuxFile#1{\message{closing file: #1. }%    % plus other actions    }\CloseAuxFiles<Revwhich gives the output\Ver>closing file: toc.  closing file: lof. <Rev\endinput\spoint Dynamic macro definitionUnpacking token lists inside an \cs{edef} can be put to arather ambitious use: dynamic definition of macros.Consider a simple example.\alttWe set ourselves the goal of lettingthe user define macros, without ever having to use \cs{def}.The syntax for this could look like\Ver>\startdefinition\do:this\do:that\define:MyMacro<Revsuch that \ver>\show\MyMacro> gives \Ver>> \MyMacro=macro:->\this \that .<RevAn implementation of this uses a token list to collectthe commands that the user specifies:\Ver>\newtoks\actionlist<RevThe first command is easy:\Ver>\def\startdefinition{\actionlist{}}<RevNow the \cs{do} command has to hang control sequencesin the \cs{actionlist}:\Ver>\def\do:#1 {%    \edef\act{\noexpand\appendaction              \expandafter\noexpand\csname#1\endcsname}%    \act}<RevThe \cs{edef} is used solely to form the actual control sequence.The next macro uses \cs{edef} to unpack the \cs{actionlist} so far:\Ver>\def\appendaction#1{%    \edef\act{\noexpand\actionlist=                 {\the\actionlist \noexpand#1}}%    \act}<RevFinally, definition of the user macro also needs an \cs{edef}.Some \cs{expandafter} trickery is necessary here to formthe control sequence of the user macro:\Ver>\def\define:#1 {%    \expandafter\edef\csname#1\endcsname{\the\actionlist}}<RevOf course, this is a very simple, rather pointless, example.However, it illustrates an important principle of howtoken lists can be used to implement another syntax levelin \TeX\ (see~\cite[EL]). This principle underlies the\term Lollipop\par`Lollipop' format that was used to typeset this book.\endinput