\subject[trace]  Tracing\TeX's workings are often quite different from what\term tracing\par\term statistics\parthe programmer expected, so there are ways to discover how \TeX\arrived at the result it did. The \cs{tracing...} commands writeall information of a certain kind to the log file (and to the terminal if \cs{tracingonline} is positive),and a number of \cs{show...} commands can be used to ask thecurrent status or value of various items of \TeX.In the following list, only \cs{show} and \cs{showthe}display their output on the terminal by default,other \cs{show...} and \cs{tracing...} commandswrite to the log file. They will write in addition tothe terminal if \cs{tracingonline} is positive.\invent\item meaning       Give the meaning of a control sequence as a string of characters.\item show       Display the meaning of a control sequence.\item showthe       Display the result of prefixing a token with \cs{the}.\item showbox       Display the contents of a box.\item showlists       Display      the contents of the partial lists      currently built in all modes.      This is treated on      page~\pgref[showlists].\item tracingcommands      If this is~1 \TeX\ displays primitive commands executed;       if this is 2~or more the outcome of conditionals is also recorded.      \csterm tracingcommands\par\item tracingmacros       If this is~1, \TeX\ shows expansion of macros       that are performed and the actual values of the arguments;       if this is 2~or more \gr{token parameter}s such as      \cs{output} and \cs{everypar} are also traced.      \csterm tracingmacros\par\item tracingoutput       If this is positive, the log file shows a dump of boxes       that are shipped to the \n{dvi} file.\item showboxdepth        The number of levels of box dump that are shown when       boxes are displayed.\item showboxbreadth       Number of successive elements on each level that are shown when       boxes are displayed.\item tracingonline       If this parameter is positive, \TeX\ will write trace            information to the terminal in addition to the log file.      \csterm tracingonline\par\item tracingparagraphs       If this parameter is positive, \TeX\ generates            a trace of the line breaking algorithm.\item tracingpages       If this parameter is positive, \TeX\ generates            a trace of the page breaking algorithm.      \csterm tracingpages\par\item tracinglostchars       If this parameter is positive, \TeX\ gives            diagnostic messages whenever a character is accessed that            is not present in a font.      \csterm tracinglostchars\par      Plain default:~1.\item tracingrestores       If this parameter is positive, \TeX\ will report            all values that are restored when a group ends.      \csterm tracingrestores\par\item tracingstats       If this parameter is~1, \TeX\ reports at the            end of the job the usage of various internal arrays;      if it is~2, the memory demands are given whenever      a page is shipped out.\inventstop\point Meaning and content: \cs{\show}, \cs{\showthe}, \cs{\meaning}The meaning of control sequences, and the contents of thosethat represent internal quantities, can be obtainedby the primitive commands \cs{show}, \cs{showthe},and~\cs{meaning}.The control sequences \cs{show} and \cs{meaning} are similar:\altthe former will give\csterm show\par\csterm meaning\paroutput to the log file and the terminal, whereas the latterwill produce the same tokens, but they are placed in \TeX'sinput stream.The meaning of a primitive command of \TeX\ is that command itself:\Ver>\show\baselineskip<Rev gives\Ver>\baselineskip=\baselineskip<RevThe meaning of a defined  quantity is its definition:\Ver>\show\pageno<Rev gives\Ver>\pageno=\count0<RevThe meaning of a macro is its parameter text and replacement text:\Ver>\def\foo#1?#2\par{\set{#1!}\set{#2?}}\show\foo<Rev gives\Ver>\foo=macro:#1?#2\par ->\set {#1!}\set {#2?}<RevFor macros without parameters the part before the arrow(the parameter text) is empty.The \cs{showthe} command will display on the log file and terminal \csterm showthe\parthe tokens that \cs{the} produces. After \cs{show}, \cs{showthe}, \cs{showbox}, and \cs{showlists}\TeX\ asks the user for input; this can be preventedby specifying \cs{scrollmode}.Characters generatedby \cs{meaning} and \cs{the} have category~12, except for spaces(see page~\pgref[cat12]);the value of \cs{escapechar} is used when control sequencesare represented.\point Show boxes: \cs{\showbox}, \cs{\tracingoutput}If \cs{tracingoutput} is positive the log file will\csterm tracingoutput\par\csterm showbox\parreceive a dumped representation of all boxes that arewritten to the \n{dvi} file with \cs{shipout}.The same representation is usedby the command \cs{showbox}\gr{8-bit number}.In the first case \TeX\ will report `Completed box being shipped out';in the second case it will enter \cs{errorstopmode}, andtell the user `OK. (see the transcript file)'.If \cs{tracingonline} is positive, the box is also displayedon the terminal; if \cs{scrollmode} has been specified,\TeX\ does not stop for input.The upper bound on thenumber of nested boxes that is dumped is \cs{showboxdepth};\csterm showboxdepth\par\csterm showboxbreadth\pareach time a level is visited at most \cs{showboxbreadth}items are shown, the remainder of the list is summarizedwith~\n{etc.}For each box its height, depth, and widthare indicated in that order, and for characters it isstated from what font they were taken. \example After\Ver>\font\tenroman=cmr10 \tenroman\setbox0=\hbox{g}\showbox0<Rev the log file will show\Ver>\hbox(4.30554+1.94444)x5.00002.\tenroman g<Rev indicating that the box was \n{4.30554pt} high,\n{1.94444pt} deep, and \n{5.00002pt} wide, and that it containeda character `g' from the font \cs{tenroman}. Note that the fifth decimal of all sizes may be roundedbecause \TeX\ works with multiples of $2^{-16}$\n{pt}.\message{ifmath: scriptfont fam0 fill!}\>The next example has nested boxes, \Ver>\vbox{\hbox{g}\hbox{o}}<Revand it contains \cs{baselineskip} glue between the boxes.After a \cs{showbox} command the log file output is:\Ver>\vbox(16.30554+0.0)x5.00002.\hbox(4.30554+1.94444)x5.00002..\tenroman g.\glue(\baselineskip) 5.75002.\hbox(4.30554+0.0)x5.00002..\tenroman o<RevEach time a new level is entered an extra dot is added tothe front of the line. Note that \TeX\ tells explicitlythat the glue is \cs{baselineskip} glue;it inserts names like this for all automatically inserted glue.The value ofthe baselineskip glue here is such that the baselines ofthe boxes are at 12 point distance.Now let us look at explicit (user) glue. \TeX\ indicates the ratioby which it is stretched or shrunk. \examples\Ver>\hbox to 20pt {\kern10pt \hskip0pt plus 5pt}<Revgives (indicating that the available stretch has been multiplied by~\n{2.0}):\Ver>\hbox(0.0+0.0)x20.0, glue set 2.0.\kern 10.0.\glue 0.0 plus 5.0<Revand\Ver>\hbox to 0pt {\kern10pt \hskip0pt minus 20pt}<Revgives (the shrink has been multiplied by~\n{0.5})\Ver>\hbox(0.0+0.0)x0.0, glue set - 0.5.\kern 10.0.\glue 0.0 minus 20.0<Revrespectively.\>This is an example with infinitely stretchable or shrinkableglue:\Ver>\hbox(4.00000+0.14000)x15.0, glue set 9.00000fil<RevThis means that the horizontal box contained \n{fil} glue, andit was set such that its resulting width was \n{9pt}.Underfull boxes are dumped like all other boxes, butthe usual `\n{Underfull hbox detected at line...}'is given. Overfull horizontal boxes contain a vertical ruleof width \cs{overfullrule}:\Ver>\hbox to 5pt {\kern10pt}<Rev gives\Ver>\hbox(0.0+0.0)x5.0.\kern 10.0.\rule(*+*)x5.0<RevBox leaders are not dumped completely:\Ver>.\leaders 40.0..\hbox(4.77313+0.14581)x15.0, glue set 9.76852fil...\tenrm a...\glue 0.0 plus 1.0fil<Revis the dump for\Ver>\leaders\hbox to 15pt{\tenrm a\hfil}\hskip 40pt<RevPreceding or trailing glue around the leaderboxes is also not indicated.\point Global statisticsThe parameter \cs{tracingstats} can be used to force \TeX\\csterm tracingstats\parto report at the end of the job the global use of resources.Some production versions of \TeX\ may not have this option.As an example, here are the statistics for this book:\Ver>Here is how much of TeX's memory you used:<RevString memory (bounded by `pool size'):\Ver> 877 strings out of 4649 9928 string characters out of 61781<RevMain memory, control sequences, font memory:\Ver> 53071 words of memory out of 262141 2528 multiletter control sequences out of 9500 20137 words of font info for 70 fonts,             out of 72000 for 255<RevHyphenation:\Ver> 14 hyphenation exceptions out of 607<RevStacks: input, nest, parameter, buffer, and save stack respectively,\Ver> 17i,6n,19p,245b,422s stack positions out of  300i,40n,60p,3000b,4000s<Rev\endinput\point Line breaking: \cs{\tracingparagraphs}If \cs{tracingparagraphs} is positive, \TeX's line breaking\csterm tracingparagraphs\paralgorithm will generate trace output. However, on some \TeX\implementations this trace mode may have been disabled to get a faster running system.Consider an example paragraph of \TeX:\Ver>\hsize=3in \parindent=0cm \frenchspacing\pretolerance=500This is a sample paragraph to show the trace output that\TeX's line breaking algorithm produces. Some \TeX\ systemscannot generate this trace, as the relevant piece of codehas been commented out for speed optimisation. With ever faster computers this won't be necessary any more.<Rev\TeX\ first attempts to break the paragraph withouthyphenation, and it will accept solutions where eachline has a badness less than \cs{pretolerance}.\Ver>@firstpass<Rev Report that the first pass has started;\Ver>[]\tenrm This is a sample paragraph to show the trace <RevApparently this is the only way to fill the first line;\Ver>@ via @@0 b=263 p=0 d=84529<Rev and doing sohad badness~263, a zero penalty, and a resulting 84529demerit points.\Ver>@@1: line 1.0 t=84529 -> @@0<Rev Conclusion:breakpoint~1 (\ver>@@1>) occurs on line~1, and it makes theline `very loose' (indicated by the~\n{.0}), and the total demerits are84529 if the previous breakpoint was number~0.The first pass is now aborted.\Ver>@secondpass[]\tenrm This is a sam-ple para-graph to show the trace out-@\discretionary via @@0 b=2 p=50 d=2644@@1: line 1.2- t=2644 -> @@0<RevWith a very small badness of~2, but with 50 penalty pointsfor breaking at a hyphen, this line is `decent' (indicated by the~\n{.2}), and the total of demerit pointsis~2644.The second line is also straighforward:\Ver>put that T[]X's line break-ing al-go-rithm pro-duces. @ via @@1 b=0 p=0 d=100@@2: line 2.2 t=2744 -> @@1<RevThe demerits now derive solely from the \cs{linepenalty},which is~10. Similarly the third line:\Ver>Some T[]X sys-tems can-not gen-er-ate this trace, as @ via @@2 b=1 p=0 d=121@@3: line 3.2 t=2865 -> @@2<RevFor the fourth line two possibilities exist:it can be set `loose' with 9409 demerit points\Ver>the rel-e-vant piece of code has been com-mented @ via @@3 b=87 p=0 d=9409@@4: line 4.1 t=12274 -> @@3<Revor, fitting in an extra word, it can be set `tight' with2601 demerit points\Ver>out @ via @@3 b=41 p=0 d=2601@@5: line 4.3 t=5466 -> @@3<RevLine 5 can be set in three ways:coming from breakpoint~4 it can be broken as\Ver>for speed op-ti-mi-sa-tion. With ever faster com-@\discretionary via @@4 b=0 p=50 d=2600@@6: line 5.2- t=14874 -> @@4<Revand coming from breakpoint~5 there are two ways:\Ver>put-@\discretionary via @@5 b=2 p=50 d=2644@@7: line 5.2- t=8110 -> @@5<Revand \Ver>ers @ via @@5 b=84 p=0 d=8836@@8: line 5.3 t=14302 -> @@5<RevOf the three, the last possibility is the only one thatdoes not involve hyphenating line~5.As line 6 is the last line of the paragraph, coming frombreakpoints 6 or~7 gives an extra 5000 demerit pointsfrom the \cs{finalhyphendemerits}.\Ver>this won't be nec-es-sary any more. @\par via @@6 b=0 p=-10000 d=5100@\par via @@7 b=0 p=-10000 d=5100@\par via @@8 b=0 p=-10000 d=100@@9: line 6.2- t=13210 -> @@7<RevHowever, coming from breakpoint 7 still gives the leastdemerits.